The complex network of trophic interactions in a subAntarctic Marine Protected Area

### 2023-03-28 MS_ToSubmit/MS_Burdwood_foodweb.tex
Add LS & IR minor comments

  \journal{MEPS, Oikos, Polar Biology (?)} % Sets Journal name

@@

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \cortext[cor1]{Corresponding author}

@@

Introduction to MPAs and MPA N-BB

Ecological importance of MPA N-BB to the region

Trophic ecology \& keystone spp knowledge of B Burdwood

Why using a network approach

In the present work we present the first, detailed analysis of the food
web for the MPA Namuncurá - Banco Burdwood ecosystem. For this we
applied a network approach to a highly-resolved food web. The objective
was twofold: characterise the food web in terms of complexity and
structure, and describe the species' role in such network framework.

@@

The MPA Namuncurá - Banco Burdwood is a shallow submarine plateau called
Burdwood Bank (BB) located 150 km east of Isla de los Estados and 200 km
south from Malvinas/Falkland Islands (Figure 1). It comprises nearly
34.000 km\textsuperscript{2} circumscribed by the 200 m isobath, between
54º--55ºS and 56º--62ºW, with a slight slope extended nearly 370 km
east--west. The BB is surrounded by steep flanks of more than 3000 m
depth through which strong currents circulate (Matano, Palma, \& Combes,
2019; A. R. Piola \& Gordon, 1989; Reta, 2014). Intense upwelling and
mixing occur over it, entraining deep nutrient rich waters into the
photic layer (Matano et al., 2019; A. Piola \& Falabella, 2009), and
resulting in a fairly homogeneous water column both spatially and
temporally (Glorioso \& Flather, 1995; Guerrero, Baldoni, \& Benavides,
1999; Matano et al., 2019). Physical features in the MPAN-BB are fairly
stable, with salinity averaging 34 all year round and temperature
ranging between 4 and 8ºC overall (Acha, Mianzan, Guerrero, Favero, \&
Bava, 2004; Guerrero et al., 1999; A. Piola \& Falabella, 2009).

\begin{figure}
\includegraphics[width=1\linewidth]{MPABurdwood_map} \caption{Map of the Marine Protected Area Namuncurá - Banco Burdwood. Taken from https://www.argentina.gob.ar/parquesnacionales/areasmarinas/namuncura-burdwood.}\label{fig:figure1}
\end{figure}

@@

In order to build the network of predator-prey interactions we reviewed
more than 150 references considering published articles, databases and
doctoral theses. Furthermore, we took into account personal
communications from experts belonging to the working group of the study
area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of
the expertise of the authors contributing to the present study was a key
factor in enhancing the quality of the network, and inherently improved
the network representation of the MPA Namuncurá-Banco Burdwood
ecosystem. A list of the references used to build the network is
presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
MPA Namuncurá-Banco Burdwood, we followed the concept of trophic
species, here defined as aggregated groups of taxa. In most cases we
followed this when data on specific biological species were not
available, but for some cases we collapsed species when taxa shared the
same set of predators and prey (trophic similarity criteria). Details
about this can be found in Supplementary Material (Table S2).

With the gathered trophic data we constructed an interaction matrix of
pairwise interactions; a value of 1 or 0 was assigned to each element
a\_ij of the matrix depending on whether the j-species preyed or not on
the i-species. Then we transformed such matrix into an oriented graph
with L trophic interactions between S nodes or species. The orientation
or direction of the graph follows the flow of energy and matter in the
network, from prey to predator.

@@

We analysed the MPA Namuncurá - Banco Burdwood network of trophic
interactions, hereafter food web, at two levels: 1) network, considering
species and interactions of the whole network; and 2) species,
considering interactions and species related to a particular species.
The network-level analysis aims to characterize the network of trophic
interactions in terms of complexity and structure. For this we
calculated several network properties commonly used to describe
empirical food webs (Pascual \& Dunne, 2005): (1) number of species S;
(2) number of interactions or links L; (3) link density L/S; (4)
Connectance L/S\^{}2; (5) omnivory Omn; and (6) shortest path length
SPL. Also, we estimated the (7) degree distributions for the food web,
for prey and predators, and for each functional group (e.g.~Amphipoda,
Ascidiacea, Bivalvia, Fish, Marine mammals, Sea birds, etc.) (Table 1).
The prey and predator distributions indicate the frequency of prey among
predators, and viceversa; the functional group's degree shows the
distribution of interactions within groups. The species-level analysis
aims to describe the species' role in the food web. For this we
considered the following properties: (1) betweenness Btw; (2) closeness
Cl; (3) trophic similarity TS; (4) topological role TR; and (5) trophic
level TL (Table 1). We also studied the relationship between species
trophic level and the other species properties by performing linear
regression analyses. Thus, we considered the IS as the dependent
variable and the given unweighted property as the independent variable,
and obtained the coefficients (slope and intercept) for the linear
model. Models were fitted using the least squares approach. We also
explored the mean IS distribution with the species habitat. These
properties provide a general appropriate description of species' role in
empirical complex food webs (Cirtwill et al., 2018).

@@

are actually realized. & It measures the probability of interactions; it
is a fundamental measure of network complexity. Connectance can be
negatively or positively associated with food web robustness, depending
on the network structure (random vs non-random) or how the strength of
the interactions are distributed. & Martinez 1992 \\

@@

\textbf{Shortest path length} & Average shortest path length between all
pairs of species. & A short path length imply a rapid spread of an
impact (e.g.~invasion, population fluctuation, local extinction)
throughout the food web. & Watts \& Strogatz 1998 \\

@@

connector. Species with the same role are expected to have similar
topological properties. & Guimera \& Nunes Amaral 2005 \\

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web
consisted of 1778 predator-prey interactions and 379 species, where 93\%
of them were defined at the species taxonomical level (Figure 2, Table
S2). The food web presented a link density of 4.69, meaning the average
number of interactions per species, and a connectance of 0.01. Almost
half of the consumers were omnivores, feeding on sources at different
trophic levels. The food web showed a path length of 2.99, which implies
that nearly three interactions are needed to connect any pair of species
of the network (Table 1).

@@

\caption{Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicate the trophic level.}\label{fig:figure2}
\end{figure}

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{Path length}\\
\hline
379 & 1778 & 4.69 & 0.01 & 0.48 & 2.99\\
\hline
\end{tabular}
\end{table}

@@

\caption{Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level; groups with less than 3 species were not plotted (e.g. pelagic fish).}\label{fig:figure3}
\end{figure}

@@

The majority of the species of the food web were consumers, 336 out of
379; the rest were primary producers, such as diatoms (phytoplankton),
and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). In
this regard, the most evident significant relationship was with trophic
similarity, i.e.~the higher the species' TL, the lower the trophic
similarity or the higher the uniqueness in terms of trophic role (Figure
4C). Here it's noteworthy to highlight high-trophic level species with
low values of trophic similarity: \emph{Bathyraja macloviana} and
\emph{Squalus acanthias} (Chondrichthyans), \emph{Diplopteraster clarki}
and \emph{Pteraster sp} (echinoderms), \emph{Phalacrocorax\_atriceps}
and \emph{Eudyptes chrysocome} (sea birds), and \emph{Lagenorhynchus
cruciger} and \emph{Mesoplodon bowdoini} (marine mammals) (Table S3).

We also found a negative significant relationship with closeness,
however less evident, meaning that low-TL species are relatively closer
to any other species in the food web (Figure 4B). Species of genera
\emph{Calanus} and \emph{Euphausia}, and species of Brachiopoda, all of
them with TL \textless{} 3, registered the highest values of closeness
(Table S3).

It's noteworthy that the highest values of betweenness were shown by
species of mid-TLs (3-4), meaning that those species participated in the
highest number of shortest paths between species (Figure 4A). These were
the species with the highest values: \emph{Patagonotothen ramsayi},
\emph{Dissostichus eleginoides}, \emph{Salilota australis} (fishes),
\emph{Doryteuthis gahi} (Cephalopoda), and \emph{Patagonotothen
guntheri} (Notothenioid fish) (Table S3).

Taking into account the topological role, `module specialist' species
were the most frequent and presented a wide TL range (1 - 4.77); `module
hub' was constrained to mid-TL species (2.48 - 3.92); `module connector'
from low to mid-TLs (1 - 3.86); and `network connector', the least
frequent, had all of its species in TL = 1, except for one with TL =
3.47 (Figure 4D, see Figure S2 for species' topological roles in a food
web graph framework). Here it's important to highlight species of the
two latter topological roles, because they are responsible for linking
modules and maintaining the connectivity of the food web: 40 species (5
network connectors + 35 module connectors) from 20 different functional
groups with a TL range = 1 - 3.86.

@@

\hypertarget{the-mpa-namuncuruxe1---banco-burdwood-food-web}{%
\subsection{The MPA Namuncurá - Banco Burdwood food
web}\label{the-mpa-namuncuruxe1---banco-burdwood-food-web}}

\begin{itemize}
\item
  The MPA Namuncurá - Banco Burdwood food web is one of the most
  highly-resolved networks of trophic interactions described for
  subpolar and polar regions. Compare network-level properties with
  other subpolar and polar food webs.
\item
  The food web shows an asymmetry in the distribution of interactions at
  several levels: food web, functional group, prey and predators. In all
  of them interactions are concentrated in a few species. On one hand,
  this asymmetry indicates that most consumers have a narrow diet, and
  few present flexible diets (\emph{Patagonotothen guntheri}, \emph{P.
  ramsayi}, \emph{Dissostichus eleginoides}, \emph{Bathyraja
  brachyurops}, \emph{B. griseocauda}). On the other hand few prey are
  consumed by many predators, meaning that there are dominant food
  sources from which most consumers depend on: detritus, diatoms
  (benthic, centric, pennate), and species of \emph{Euphausia}.
\end{itemize}

\hypertarget{species-role}{%
\subsection{Species' role}\label{species-role}}

Our results state that certain species play particular roles in the
structure of the food web.

\begin{itemize}
\item
  Betweenness. Species at mid-trophic levels (3-4) lie in more shortest
  paths between species than any other species: `bridge' role.
  Implications for spread of perturbations like mercury (Fioramonti,
  Ribeiro Guevara, Becker, \& Riccialdelli, 2022), microplastics (Cossi
  et al., 2021).
\item
  Closeness. Low-trophic level species are relatively closer to other
  species in the food web. Apart from the most demanded food sources
  (i.e.~detritus), which are logically closer to many species,
  unexpected consumers arise as important in this regard. Species of the
  zooplankton community like \emph{Calanus} and \emph{Euphausia} are
  crucial for the food web functioning, since they can affect others
  more quickly.
\item
  Trophic similarity. Species are more similar at low and mid-trophic
  levels (2-4). There is a functional redundancy at these levels of the
  food web. On the other hand, less trophic similarity at high TLs
  indicate uniqueness in terms of trophic role. This highlights the
  importance of top predators in the sense that they play a role that
  cannot be played by any other species.
\item
  Topological role. Low and mid-trophic level species (TL = 1 - 3.86)
  are responsible for maintaining the connectivity of the food web
  (network connector and module connector). The functional groups
  represented are: Amphipoda, Bivalvia, Bryozoa, Cnidaria, Cumacea,
  Decapoda, Detritus, Diatoms, Echinodermata, Fish (bentho-pelagic,
  demersal, chondrichthyes), Foraminifera, Polychaeta, Porifera,
  Zooplankton. This shows that the structure of the food web depends on
  a diversity of species, which at the end comprise the ecosystem's
  diversity.
\item
  Overall, low and mid-trophic level species arise as crucial in the
  structure of the MPA Namuncurá - Banco Burdwood food web, since they
  present the highest betweenness and closeness values, and the most
  important topological roles. This coincides with a recent study that
  suggest a wasp-waist trophic structure for the region (Riccialdelli et
  al., 2020). However, we also suggest that there exists functional
  redundancy at mid-trophic levels. Furthermore, our study points out
  species of high-trophic levels due to its uniqueness role.
\end{itemize}

>>>>

  \journal{MEPS, Oikos, Ecological Applications, J of Applied Ecology
(?)} % Sets Journal name

@@

    \author[]{To be continued(?)\textsuperscript{c}%
  %
  }

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \affiliation[3]{INIDEP, UBA-CONICET, IIMyC-CONICET, MACN-CONICET,
UNC-CONICET}
    \cortext[cor1]{Corresponding author}

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to nearly
29,600,000 km\textsuperscript{2}, distributed across nearly 18,444 MPAs
and covering 8.16\% of the ocean's surface (IUCN \& UNEP-WCMC, 2023),
and therefore approaching the 10\% goal of the Convention of Biological
Diversity (CBD, 2004). Despite this progress, recent reports have shown
that actual protection has been overestimated because it includes areas
that are not yet effectively protected (only declared) as well as areas
that allow significant extractive activities (Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017). Added to
this, such MPAs are part of a network of protected areas in the
sub-Antarctic (jointly with MPA Yaganes), that aims to protect and
strengthen the functioning of this southern region to maintain global
ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, the presence of large numbers (or
proportions) of endemic species, and/or the presence of high
biodiversity across taxonomic levels (Hogg et al., 2016). Indeed, the
MPA N-BB was created to protect a sensitive and biodiverse benthic
habitat that was barely known (Falabella, 2017; L. Schejter et al.,
2016). The benthic community is featured by a high biomass of vulnerable
and fragile species (mainly Poriferaand Cnidaria) that meet the
characteristics of vulnerable marine ecosystems (Schejter \& Albano,
2021), here defined as sites that present taxa with a minimum biomass of
10 kg per 1200 m\textsuperscript{2} (Commission for the Conservation of
Antarctic Marine Living Resources, 2009). Also, it provides habitat to
several small-sized species (López-Gappa et al., 2018; L. Schejter \&
Bremec, 2019; Sirito, 2019), and has an important role in the life
history of fishes as a food source, refuge and nursery area (Delpiani et
al., 2020; García Alonso et al., 2018; Troccoli et al., 2020). The
maintenance of this singular community is related to local and regional
oceanographic processes, including the circulation of the rich Malvinas
(Falkland) current in the area (Guerrero et al., 1999; A. R. Piola \&
Gordon, 1989) and the upwelling and mixing phenomena (Matano et al.,
2019). The input of nutrients coming from the Malvinas (Falkland)
current also supports a diverse plankton community (Guinder et al.,
2020). Overall, 811 benthic and plankton species have been identified
for the MPA N-BB ecosystem, where 349 were reported for the first time
for the area (APN, 2022). However, identifying the main species for
maintenance of ecosystem services and health as well as for management
and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem is regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure than other sub-Antarctic areas, such as the Atlantic coast of
Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis} and
longtail southern cod \emph{Patagonotothen ramsayi} are considered the
most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are characterized
by an unexpected complexity in terms of biodiversity and ecological
interactions (Cordone et al., 2020; Day et al., 2013; Kortsch et al.,
2019; Trathan et al., 2021). Although there is a robust knowledge about
the complexity considering the richness of the benthic and plankton
communities in the MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020;
L. Schejter et al., 2016, 2020), there is a need for a better
understanding of species interactions' complexity and structure. This
aspect can be tackled by analyzing one of the most-frequent
relationships occurring between species: the predator-prey interaction
(Bascompte, 2009). The sum of predator-prey or trophic interactions of a
particular region is referred to as a food web, which represents the
roadmap for matter and energy flow in an ecosystem. In recent years,
network approaches have been successfully applied to study complex
high-latitude marine ecosystems, improving our knowledge on structure,
functioning, and response to environmental/anthropogenic changes
(Cordone et al., 2018; Funes et al., 2022; Kortsch et al., 2015; Marina
et al., 2023). Among anthropogenic threats, it is worth mentioning that
contaminants like mercury and microplastics have been recently reported
as important threats for the MPA N-BB region (Cossi et al., 2021; Di
Mauro et al., 2022; Fioramonti et al., 2022).

In the present work, we present the first, detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterize the food web
in terms of complexity and structure, and identify the species' role in
such network framework.

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023). They are located 150 km east of isla de
los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km\textsuperscript{2}
circumscribed by the 200 m isobath, between 54º--55ºS and 56º--62ºW,
with a slight slope extended nearly 370 km east--west. Physical features
in the BB are fairly stable, with salinity averaging 34 all year round
and temperature ranging between 4 and 8ºC (Acha et al., 2004; Guerrero
et al., 1999; A. Piola \& Falabella, 2009). The BB is surrounded by
steep flanks of up to 4000 m depth through which strong currents
circulate (Matano et al., 2019; A. R. Piola \& Gordon, 1989; Reta,
2014). Such deep slope form the basis of N-BB II that protects about
32,000 km\textsuperscript{2}. Intense upwelling and mixing occur over
it, entraining deep nutrient-rich waters into the photic layer (Matano
et al., 2019; A. Piola \& Falabella, 2009), and resulting in a fairly
homogeneous water column both spatially and temporally (Glorioso \&
Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in APN, 2022), a joint management plan was recently proposed
(APN, 2022).

\begin{figure}
\includegraphics[width=1\linewidth]{MPABurdwood_map} \caption{Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve.}\label{fig:figure1}
\end{figure}

@@

In order to build the network of predator-prey interactions we reviewed
more than 150 references considering published articles, databases and
doctoral theses. Furthermore, we took into account personal
communications from experts belonging to the working group of the study
area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of
the expertise of the authors contributing to the present study was a key
factor in enhancing the quality of the network, and inherently improved
the network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserves food web functional properties (Gauzens et
al., 2013). Details about this can be found in Supplementary Material
(Table S2).

With the gathered trophic data we constructed a matrix of pairwise
interactions; a value of 1 or 0 was assigned to each element a\_ij of
the matrix depending on whether the j-species preyed or not on the
i-species. Then we transformed such matrix into an oriented graph with L
trophic interactions between S nodes or species. The orientation or
direction of the graph follows the flow of energy and matter in the
network, from prey to predator.

@@

We analyzed the MPA N-BB network of trophic interactions, or food web,
at two levels: A) network, considering species and interactions of the
whole network; and B) species, considering interactions and species
related to a particular species (Table 1).

The network-level analysis aims to characterizee the food web in terms
of complexity and structure. For this we calculated several network
properties commonly used to describe empirical food webs (Pascual \&
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon we analyzed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is defined as the average shortest
path length between all pairs of nodes (Watts \& Strogatz, 1998). Here,
CPL was calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

\[
CPL = \frac{2}{S(S-1)} \sum_{i = 1}^{S} \sum_{i = 1}^{S} {CPL_{Min}(i,j)}
\] The CC quantifies the local interconnectedness of the network and it
is defined as the fraction of the number of existing links between
neighbours of node \(i\) among all possible links between these
neighbours. In this study, the CC was determined as the average of the
individual clustering coefficients \(CC_i\) of all the nodes in the
network. Individual \(CC_i\) were determined as follows:

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). With the aim of testing whether the
food web presented the small-world pattern, we compared the empirical
values of CPL and CC with those resulting from 1000 randomly generated
networks with the same size (S) and number of interactions (L),
following the method proposed by Marina et al. (2018).

Also, we estimated the (7) degree distributions for the food web, for
prey and predators, and for each functional group (e.g., Amphipoda,
Ascidiacea, Bivalvia, Fish, Marine mammals, Sea birds, etc.). The prey
and predator distributions indicate the frequency of prey among
predators, and viceversa; the functional group's degree shows the
distribution of interactions within groups.

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: (1)
betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4)
topological role TR; and (5) trophic level TL (Table 1). Topological
roles refers to the fact that food webs tend to naturally organize in
non-random, modular patterns, where modules are defined as a group of
species that interact more frequently among themselves than with species
that are not members of the module (Guimerà \& Nunes Amaral, 2005).
Species can play different roles in this respect, according to the
pattern of interactions they have within their own module and/or across
modules. We computed the topological role for each species, classified
as: module hub, species with a relatively high number of interactions,
but most within its own module; module specialist, species with
relatively few interactions and most within its own module; module
connector, species with relatively few interactions mainly between
modules; and network connector, species with high connectivity between
and within modules (Guimerà \& Nunes Amaral, 2005).

We also studied the relationship between species TL and the other
species properties by performing linear regression analyses. Thus, we
considered the TL as the dependent variable and the given property
(i.e.~betweenness, closeness, trophic similarity) as the independent
variable, and obtained the coefficients (slope and intercept) for the
linear model. Models were fitted using the least squares approach. We
also explored the topological role categories with the species TL. These
species-level properties provide a general appropriate description of
species' role in empirical complex food webs (Cirtwill et al., 2018).

@@

are actually realized. Range = 0 - 1. & It measures the probability of
interactions and is a fundamental measure of network complexity.
Connectance can be negatively or positively associated with food web
robustness, depending on the network structure (random vs non-random) or
how the strength of the interactions are distributed. & Martinez 1992 \\

@@

\textbf{Small-world pattern} & A network with short path length
(distance between nodes) and high clustering coefficient (formation of
compartments) compared to random networks. & Consequences of this
structural pattern in food webs are of great importance in recognizing
evolutionary paths and the vulnerability to perturbations. & Watts \&
Strogatz 1998, Montoya \& Solé 2002 \\

@@

connector. Network connector and module connector roles maintain the
connectivity of the food web. & Guimera \& Nunes Amaral 2005 \\

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web
(N-BB I and II) consisted of 1778 predator-prey interactions and 379
species, where 93\% of them were defined at the species taxonomical
level (Figure 2, Table S2). The food web presented a link density (e.g.,
the average number of interactions per species) of 4.69, and a
connectance of 0.01. Almost half of the consumers were omnivores (0.48),
feeding on sources at different trophic levels. The food web displayed a
small-world pattern, meaning that the path length was lower and the
clustering coefficient higher than the random networks (Table 2).

@@

\caption{Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level.}\label{fig:figure2}
\end{figure}

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
379 & 1778 & 4.69 & 0.01 & 0.48 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

@@

\caption{Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level; groups with less than 3 species were not plotted (e.g., pelagic fish).}\label{fig:figure3}
\end{figure}

@@

The majority of the species in the food web were consumers, 336 out of
379; the rest were primary producers, such as diatoms (phytoplankton),
and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species with low values
of trophic similarity: \emph{Bathyraja macloviana} and \emph{Squalus
acanthias} (Chondrichthyans), \emph{Diplopteraster clarki} and
\emph{Pteraster sp} (echinoderms), \emph{Phalacrocorax\_atriceps} and
\emph{Eudyptes chrysocome} (sea birds), and \emph{Lagenorhynchus
cruciger} and \emph{Mesoplodon bowdoini} (marine mammals) (Table S3).

We also found a negative significant relationship between TL and
closeness, however less evident, meaning that low-TL species are
relatively closer to any other species in the food web (Figure 4B).
Species of genera \emph{Calanus} and \emph{Euphausia}, and species of
Brachiopoda, all of them with TL \textless{} 3, registered the highest
values of closeness (Table S3).

It is noteworthy that the highest values of betweenness were shown by
species of mid-TLs (3-4), meaning that those species participated in the
highest number of shortest paths between species (Figure 4A). The
following are the species with the highest values: \emph{Patagonotothen
ramsayi}, \emph{Dissostichus eleginoides}, \emph{Salilota australis}
(fishes), \emph{Doryteuthis gahi} (Cephalopoda), and
\emph{Patagonotothen guntheri} (Notothenioid fish) (Table S3).

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.77); `module hub' was
constrained to mid-TL species (2.48 - 3.92); `module connector' from low
to mid-TLs (1 - 3.86); and `network connector', the least frequent, had
all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D,
see Figure S2 for species' topological roles in a food web graph
framework). Here it is important to highlight the two latter topological
roles, because they are responsible for linking modules and maintaining
the connectivity of the food web: 40 species (5 network connectors + 35
module connectors) from 20 different functional groups with a TL range =
1 - 3.86. The following were the five species with a network connector
role: Detritus, Diatoms (benthic, centric and pennate), and
\emph{Patagonotothen ramsayi}. The 35 species with a module connector
role represent these functional groups: Amphipoda, Bivalvia,
Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, Detritus, Diatoms,
Dinoflagellates, Echinodermata, Fish (bentho-pelagic, demersal,
chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, Sea
birds, and Zooplankton (see Supplementary Material Table S3 for the
identity of the species).

@@

\hypertarget{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}{%
\subsection{The food web of the MPA Namuncurá - Banco Burdwood
ecosystem}\label{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}}

The food web of the MPA Namuncurá - Banco Burdwood ecosystem analysed in
this study is one of the most highly-resolved networks of trophic
interactions ever studied not only for a high-latitude marine ecosystem
but also for a marine protected area worldwide. It is of paramount
importance to consider the complexity of species interactions in order
to gain insights into the structure and functioning of the ecosystem,
since the aggregation of species might mask food web properties and/or
produce type II errors (false positive) (Gauzens et al., 2013; Martinez,
1993).

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed,
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina et al., 2018). In particular, food webs from high-latitude
regions tend to exhibit a connectance closer to the minimum (between
0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al., 2022; Santana et
al., 2013). Whether food webs display a low or a high connectance helps
to better comprehend ecosystem's synthetic properties like robustness.
In this sense, empirical analyses support the notion that
highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be five times lower than that of Beagle Channel, an
adjacent area (Rodriguez et al., 2022).

The degree distribution or the distribution of number of interactions
per species is the core for the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Dunne et al.,
2002a; Montoya \& Solé, 2003; Stouffer et al., 2005), and has received
great attention in complex networks in general (Albert \& Barabási,
2002; Newman, 2003). The degree distribution affects the resilience of
complex food webs against random failures and intentional attacks: food
webs showing right-skewed distributions, like the one described in this
study, are more vulnerable to the removal of the most connected species
or hubs, with the potential of producing secondary extinctions and a
catastrophic fragmentation of the network (Albert et al., 2000; Dunne et
al., 2002b; Eklöf \& Ebenman, 2006).

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Dunne et al., 2002c; Marina et al., 2018). However, the
food web of the MPA N-BB does display a small-world pattern.
Consequences of this could be of great importance in recognizing
evolutionary paths and the vulnerability to perturbations (Montoya \&
Solé, 2002). On the one hand, a short path length implies a rapid spread
of an impact (e.g., contaminant, population fluctuation, local
extinction) throughout the network and, at the same time, a more
potentially adaptive dynamics in face of external perturbations (Montoya
\& Solé, 2002; Williams et al., 2002). On the other hand, a high
clustering coefficient indicates the formation of subnetworks composed
only by the neighbours of particular species. This translates into a
greater resistance of the network due to the confinement of
perturbations mainly within subnetworks and not spreading between them
(Heer et al., 2020; Kortsch et al., 2019). Overall, a small-world
topology provides ecological networks with a greater resilience and
resistance (Bornatowski et al., 2017; Dormann et al., 2017).

Omnivory acts as a buffer to changes as the ecosystem presents
alternative energy pathways in the face of perturbations; i.e., reducing
the risk of cascading extinctions following primary loss of species
(Borrvall et al., 2000). This is based on the fact that omnivores are
species able to adapt faster and to a wider range of environmental
conditions by changing their foraging habits to feed on the most
abundant prey (Fagan, 1997). Furthermore, omnivory can be analysed from
the interaction point of view: theoretical studies have identified
omnivorous interactions as a possible candidate for a keystone
interaction, sensu Kadoya et al. (2018), highlighting the importance of
omnivory inn stabilizing food web dynamics (McCann \& Hastings, 1997;
Neutel et al., 2002). The high proportion of omnivory in the food web of
the MPA N-BB suggests that the network might be robust to variations in
prey abundances, which could increase food web's persistence and
stability (Stouffer \& Bascompte, 2010).

To resume, the food web of the MPA Namuncurá - Banco Burdwood presents a
combination of network properties that make it unique in terms of
network resolution, complexity, and structural pattern. All this suggest
that the ecosystem might be fragile to external perturbations targeting
highly connected species. However, structural properties might provide
resilience and resistance with the final outcome of a rearrange
structure maintaining their functions.

\hypertarget{dominant-consumers-and-food-sources}{%
\subsection{Dominant consumers and food
sources}\label{dominant-consumers-and-food-sources}}

The degree distribution allows to identify important species, such as
potential keystone species (Dunne et al., 2002b; Solé \& Montoya, 2001),
generalist/specialist species, and dominant food sources (Kondoh et al.,
2010).
